The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-07 rational.tex
semantics

%subcode-inline{bnf-inline} name langRev
%! swap+ = \mathit{swap}^+
%! swap* = \mathit{swap}^*
%! dagger =  ^{\dagger}
%! assocl+ = \mathit{assocl}^+
%! assocr+ = \mathit{assocr}^+
%! assocl* = \mathit{assocl}^*
%! assocr* = \mathit{assocr}^*
%! identr* = \mathit{uniti}
%! identl* = \mathit{unite}
%! dist = \mathit{distrib}
%! factor = \mathit{factor}
%! eta = \eta
%! eps = \epsilon
%! eta+ = \eta^+
%! eps+ = \epsilon^+
%! eta* = \eta^{\times}
%! eps* = \epsilon^{\times}
%! trace+ = trace^+
%! trace* = trace^{\times}
%! ^^^ = ^{-1}
%! (o) = \fatsemi
%! (;) = \fatsemi
%! (*) = \times
%! (+) = +

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{*}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! |- = \vdash
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! langRev = \Pi
%! langRevT = \Pi^{o}
%! langRevEE = \Pi^{\eta\epsilon}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

@@

@@

>>>>

%subcode-inline{bnf-inline} name langRev
%! swap+ = \mathit{swap}^+
%! swap* = \mathit{swap}^*
%! dagger =  ^{\dagger}
%! assocl+ = \mathit{assocl}^+
%! assocr+ = \mathit{assocr}^+
%! assocl* = \mathit{assocl}^*
%! assocr* = \mathit{assocr}^*
%! identr* = \mathit{uniti}
%! identl* = \mathit{unite}
%! dist = \mathit{distrib}
%! factor = \mathit{factor}
%! eta = \eta
%! eps = \epsilon
%! eta+ = \eta^+
%! eps+ = \epsilon^+
%! eta* = \eta^{\times}
%! eps* = \epsilon^{\times}
%! trace+ = trace^+
%! trace* = trace^{\times}
%! ^^^ = ^{-1}
%! (o) = \fatsemi
%! (;) = \fatsemi
%! (*) = \times
%! (+) = +
%! LeftPlus = L^+
%! RightPlus = R^+
%! LeftTimes = L^{\times}
%! RightTimes = R^{\times}

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{*}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! |- = \vdash
%! <><> = \approx
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! langRev = \Pi
%! langRevT = \Pi^{o}
%! langRevEE = \Pi^{\eta\epsilon}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

@@

\paragraph{Primitive Isomorphsims.}
Previously the reductions of the primitive isomorphisms were specified
in the following form.

{{ iso v_{input pattern} |--> v_{output pattern} }}

\noindent
In the precense of unification, we rewrite them to accept the incoming
substitution and extend it with the appropriate unification.

{{ iso s v' |--> (v_{output pattern}, s[v' <><> v_{input pattern}]) }}

\paragraph{Extending the evaulator with unification.}
%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
% <iso, v, cc, s>_{fwd} &|-->& {[iso, v', cc, s']}_{fwd}
% & & where iso s v |--> (v', s')
% <c1(;)c2, v, cc, s>_{fwd} &|-->& <c1, v, Fst cc c2, s>_{fwd}
% {[c1, v, Fst cc c2, s]}_{fwd} &|-->& <c2, v, Snd c1 cc, s>_{fwd}
% {[c2, v, Snd c1 cc, s]}_{fwd} &|-->& {[ c1(;)c2, v, cc,s ]}_{fwd}
% <c1(+)c2, v', cc, s>_{fwd} &|-->& <c1, v, LeftPlus cc c2, s[v' <><> left v]>_{fwd}
% {[ c1, v, LeftPlus cc c2, s ]}_{fwd} &|-->& {[c1 (+) c2, left v, cc,s ]}_{fwd}
% <c1(+)c2, v', cc, s>_{fwd} &|-->& <c2, v, RightPlus c1 cc, s[v' <><> right v]>_{fwd}
% {[ c2, v, RightPlus c1 cc,s ]}_{fwd} &|-->& {[c1 (+) c2, right v, cc,s ]}_{fwd}
% <c1(*)c2, v', cc, s>_{fwd} &|-->& <c1, v1, LeftTimes cc c2 v2, s'>_{fwd}
% & & where s' = s[v <><> (v1, v2)]
% {[ c1, v1, LeftTimes cc c2 v2,s ]}_{fwd} &|-->& <c2, v2, RightTimes c1 v1 cc, s>_{fwd}
% {[ c2, v2, RightTimes c1 v1 cc,s ]}_{fwd} &|-->& {[ c1 (*) c2, (v1, v2), cc,s ]}_{fwd}

\paragraph{Backward evaluator.}
%subcode{opsem} include main
%! columnStyle = rclr
%! bck = \triangleleft
% {[iso, v, cc, s]}_{bck} &|-->& <iso, v', cc, s'>_{bck}
% & & where iso^{dagger} s v |--> (v', s')
% <c1, v, Fst cc c2, s>_{bck} &|-->& <c1(;)c2, v, cc, s>_{bck}
% <c2, v, Snd c1 cc, s>_{bck} &|-->& {[c1, v, Fst cc c2, s]}_{bck}
% {[ c1(;)c2, v, cc, s ]}_{bck} &|-->& {[c2, v, Snd c1 cc, s]}_{bck}
% <c1, v, LeftPlus cc c2,s >_{bck} &|-->& <c1(+)c2,left v, cc>_{bck}
% {[c1 (+) c2, v', cc, s ]}_{bck} &|-->& {[ c1, v, LeftPlus cc c2, s[v' <><> left v] ]}_{bck}
% <c2, v, RightPlus c1 cc, s>_{bck} &|-->& <c1(+)c2, right v, cc>_{bck}
% {[c1 (+) c2, v', cc,s ]}_{bck} &|-->& {[ c2, v, RightPlus c1 cc, s[v' <><> right v] ]}_{bck}
% <c1, v1, LeftTimes cc c2 v2, s>_{bck} &|-->& <c1(*)c2, (v1, v2), cc, s>_{bck}
% <c2, v2, RightTimes c1 v1 cc, s>_{bck} &|-->& {[ c1, v1, LeftTimes cc c2 v2, s ]}_{bck}
% {[ c1 (*) c2, v, cc, s ]}_{bck} &|-->& {[ c2, v2, RightTimes c1 v1 cc, S' ]}_{bck}
% && where s' = s[v <><> (v1, v2)]

@@

MUST TALK ABOUT AND CITE ABRAMSKY AND COECKE'S TR ABOUT QUANTUM
COMPUTING.
